Blame and Coercion: Together Again for the Second Time

### 2019-10-16 paper.tex
rm hyphen in hyper-coercion
reply
rebase on cast calculus

% Lazy D blame calculus on space-inefficient CEK
\newcommand{\judgeDBreduce}[2]{#1 \longrightarrow_{\mathcal{D}} #2}
\newcommand{\judgeDBreduceTrans}[2]{#1 \longrightarrow_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDBeval}[2]{#1 \Downarrow_{\mathcal{D}} #2}

@@

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\todoKC{rewrite this abstract based on the comparision table. -Jeremy}
  Gradual Typing is great. Implementing gradually typed with blame
  tracking and space efficiency is tricky. There exist two technique
  to do this: coercion and threesome. Coercion is easy to understand,
  and easy enough to implement, but difficult to reason about
  formally. Threesome is hard to understand, easy to implement, and
  easy to reason about formally. We propose hypercoercion, which is
  easy to understand, as easy to implement as coercion, and easy to
  reason about formally.
\todo[inline]{Spell it ``hypercoercion'' without the hyphen. -Jeremy}
\end{abstract}

@@

%\subsection{coercion-based approach and blame strategies}

\citet{siek2009exploring} decorate \citet{herman2010space}'s coercions with 
labels to restore 
blame-tracking. They also propose that there are four blame 
strategies for their coercion:
$ \{Lazy, Eager\} \times \{D, UD\} $. 
Eager strategies detect more potential type errors, 
but blame more programs.
$ D $ and $ UD $ assign blame labels differently.
They prove their \lazyUD{} coercion simulates the Blame Calculus in 
\cite{wadler2009well}. It is unknown, however, whether the simulation in the 
other direction also works.
\todo[inline]{Is it important to this paper whether there is a
  simulation in the other direction? -Jeremy}

@@

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised coercion calculi.
\todo[inline]{Not ``first''... lazy D is in the 2009 paper, right?
  Or did you mean something else? -Jeremy}
\todo[inline]{The ``Blame Calculus'' is a \lazyUD{} cast calculus.
  There is no such thing as a \lazyD{} Blame Calculus. -Jeremy}
\todo[inline]{Actually, Wadler might disagree that the blame calculus is
  a lazy UD calculus... because the blame calculus has polarity in the blame
  labels, and the lazy UD calculus does not. So perhaps we should just
  say lazy UD cast calculus, and explain that it is very close to the
  blame calculus. -Jeremy}
They simplify the coercion calculi by only 
working with normal coercions (i.e. coercions in normal form). 
They conjecture that their revised coercion calculi bisimulate the 
corresponding Blame Calculi.

@@

We prove formally that the \lazyD{} (resp. \lazyUD{}) hypercoercion calculus 
bisimulates the \lazyD{} (resp. \lazyUD{}) Blame Calculus. This is 
the first bisimulation proof for the \lazyD{} Blame Calculus as far as we know.

@@

\figref{fig:blame-static} defines the syntax of blame calculus and its
static semantics. It is little changed from previous definitions.
\todo[inline]{Which previous definitions are you refering to?  What
  are the little changes? -Jeremy} \todo[inline]{The term typing rule
  for casts is broken. -Jeremy} The dynamic semantics of Blame Calculi
depend on blame strategies, so we defer them to sub-sections.

Blame Calculus is based on Simply Typed Lambda Calculus with sum types and 
product types ($ \mathtt{STLC+} $). 
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type, $ \POOunit 
$. 
Other pre-types are functions, products, and sums.

@@

\subsection{The Blame Calculus (a \lazyD{} cast calculus)}

@@

	Reduction \fbox{$ \judgeDBreduce{s}{s} $}
	\[
	\begin{array}{rclr}

@@

	Evaluation \fbox{$\judgeDBeval{e}{o}$}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{B}^{*} 
		\sOOhalt{o}
	}{
		\judgeDBeval{e}{o}
	}
	\]

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of the \lazyD{} Blame 
Calculus in the style of CEK machines \citep{felleisen1986control}.
\todo[inline]{Briefly present the main ideas regarding how a CEK
  machine works. (2-3 sentences) -Jeremy}
\todo[inline]{Use citep to put a citation in parentheses, as I've
  done above. -Jeremy }
\todo[inline]{The figure is too tall. You may need to pull some things
  out of the figure, such as the value typing. -Jeremy}
\todo[inline]{Please fill in the missing reduction rules. -Jeremy}
\todo[inline]{Explain why you are using a CEK machine instead of
  reduction semantics. -Jeremy}
\todo[inline]{Change the order of parameters for applyCast to take
  the value first and the cast second. -Jeremy}
\todo[inline]{The definition of values is too loose for a cast.
  You need to restrict the casts to just the inert ones. -Jeremy }
We conjecture that this CEK machines agree with the \lazyD{} blame calculus in 
\citet{siek2012interpretations}.
We will show in \hyperref{sec:hypercoercion} that this abstract machine is 
equivalent to an abstract machine that represents casts with hypercoercions.

@@

Transitive closure of reduction ($ \judgeDBreduceTrans{s}{s} $) and evaluation 
are 
standard.

\subsection{The Blame Calculus (i.e. \lazyUD{} cast calculus) [Jeremy]}

The dynamic semantics of the Blame Calculus, which we define here as a
CEK machine, is similar to that of the \lazyD{} cast calculus
(Fig. \ref{machine-cekc}). The only difference is in the definition of
the $\mathit{applyCast}$ function, in which a cast whose source or
target is the unknown type $\star$ is always split into two casts that
go through an injectiable type, that is, a type in which all
sub-components are the unknown type, such as $\star \to \star$. The
definition of $\mathit{applyCast}$ for the Blame Calculus is given in
Fig.~\ref{fig:apply-Cast-UD}.

@@

\begin{lemma}[Weak Bisimulation between $ \mathcal{S}(\cdot) $ and $ 
	\mathcal{D} $]
	\label{thm:surely-monoidic-reduce}
	For all cast representations $ C $, 
	if $ C $ is surely \lazyD{} and monoidic,
	$ s_1 \in \mathcal{D} $,
	$ s_2 \in \mathcal{S}(C) $, and
	$ s_1 \approx s_2 $, 
	\begin{enumerate}
		\item If $ \judgeDBreduce{s_1}{s_3} $,
		then
		there exist $ s_5 \in \mathcal{D} $, $ s_4 \in \mathcal{S}(C) $,
		such that \begin{itemize}
			\item $ s_5 \approx s_4 $, and
			\item $ \judgeSreduce{C}{s_2}{s_4} $,
			\item $ \judgeDBreduceTrans{s_3}{s_5} $
		\end{itemize}
		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
		then
		there exist $ s_3 \in \mathcal{D} $,
		such that \begin{itemize}
			\item $ s_3 \approx s_4 $
			\item $ \judgeDBreduceTrans{s_1}{s_3} $
		\end{itemize}
	\end{enumerate}
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}

\begin{lemma}[Surely \lazyD\ and Monoidic Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-monoidic-eval}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{} and monoidic,
	\[
	\judgeDBeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{lemma}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-reduce}.
\end{proof}

@@

\begin{theorem}[\lazyD{} hypercoercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeDBeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval},
	Lemma~\ref{thm:hc-surely-lazyD}, and 
	Proposition~\ref{thm:hc-monoid}.
\end{proof}

@@

\begin{theorem}[Surely \lazyD\ Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-lazyD-correct}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{}
	\[
	\judgeDBeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}.
\end{proof}

>>>>

% Lazy D cast calculus on space-inefficient CEK
\newcommand{\judgeDCreduce}[2]{#1 \longrightarrow_{\mathcal{D}} #2}
\newcommand{\judgeDCreduceTrans}[2]{#1 \longrightarrow_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDCeval}[2]{#1 \Downarrow_{\mathcal{D}} #2}

@@

% abbrev
\newcommand{\castCalculus}[0]{$\lambda_{\rightarrow}^{\langle\cdot\rangle}$}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\todoKC{rewrite this abstract based on the comparision table. -Jeremy}
  Gradual Typing is great. Implementing gradually typed with blame
  tracking and space efficiency is tricky. There exist two technique
  to do this: coercion and threesome. Coercion is easy to understand,
  and easy enough to implement, but difficult to reason about
  formally. Threesome is hard to understand, easy to implement, and
  easy to reason about formally. We propose hypercoercion, which is
  easy to understand, as easy to implement as coercion, and easy to
  reason about formally.
\end{abstract}

@@

\citet{siek2009exploring} present another intermediate langauge for gradually 
typed language, cast calculus \castCalculus\ and achieve both blame-tracking 
and space-efficiency for this language by decorating \citet{herman2010space}'s 
coercions with labels. \castCalculus\ is simply typed lambda calculus extended 
with the dynamic type and cast expressions. The labels in \castCalculus\, 
unlike their counterparts in the Blame Calculus, have no polarity. 
They also propose that there are four blame strategies for \castCalculus{}:
$ \{Lazy, Eager\} \times \{D, UD\} $. Eager strategies detect more potential 
type errors, but blame more programs. $ D $ and $ UD $ assign blame labels 
differently. They claim that their \lazyUD\ cast calculus is the same as the 
Blame Calculus, modulo label polarity, and prove that their their \lazyUD\ 
coercion calculus simulates their \lazyUD{} cast calculus. It is unknown, 
however, whether the simulation in the other direction also works.
\todo[inline]{Is it important to this paper whether there is a
  simulation in the other direction? -Jeremy}
\todo[inline]{I guessed to say coercion is correct the other direction is 
necessary. -KC}

@@

\citet{siek2012interpretations} present a revised coercion calculi.
\todo[inline]{Actually, Wadler might disagree that the blame calculus is
  a lazy UD calculus... because the blame calculus has polarity in the blame
  labels, and the lazy UD calculus does not. So perhaps we should just
  say lazy UD cast calculus, and explain that it is very close to the
  blame calculus. -Jeremy}
\todo[inline]{I replace most uses of blame calculus to cast calculus.}
They simplify the coercion calculi by only 
working with normal coercions (i.e. coercions in normal form). 
They conjecture that their revised coercion calculi bisimulate the 
corresponding Blame Calculi.

@@

We prove formally that the \lazyD{} (resp. \lazyUD{}) hypercoercion calculus 
bisimulates the \lazyD{} (resp. \lazyUD{}) cast calculus. This is 
the first bisimulation proof for the \lazyD{} cast calculus as far as we know.

@@

\figref{fig:blame-static} defines the syntax of cast calculus and its
static semantics. It is little changed from previous definitions.
\todo[inline]{Which previous definitions are you refering to?  What
  are the little changes? -Jeremy} \todo[inline]{The term typing rule
  for casts is broken. -Jeremy} The dynamic semantics of Blame Calculi
depend on blame strategies, so we defer them to sub-sections.

Our cast calculus extend the one in \citet{siek2009exploring} with sum types 
and product types ($ \mathtt{STLC+} $). 
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type, $ \POOunit 
$. 
Other pre-types are functions, products, and sums.

@@

\subsection{The \lazyD{} cast calculus}

@@

	Reduction \fbox{$ \judgeDCreduce{s}{s} $}
	\[
	\begin{array}{rclr}

@@

	Evaluation \fbox{$\judgeDCeval{e}{o}$}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{B}^{*} 
		\sOOhalt{o}
	}{
		\judgeDCeval{e}{o}
	}
	\]

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of the \lazyD{} Blame 
Calculus in the style of CEK machines \citep{felleisen1986control}.
\todo[inline]{Briefly present the main ideas regarding how a CEK
  machine works. (2-3 sentences) -Jeremy}
\todo[inline]{Use citep to put a citation in parentheses, as I've
  done above. -Jeremy }
\todo[inline]{The figure is too tall. You may need to pull some things
  out of the figure, such as the value typing. -Jeremy}
\todo[inline]{Please fill in the missing reduction rules. -Jeremy}
\todo[inline]{Explain why you are using a CEK machine instead of
  reduction semantics. -Jeremy}
\todo[inline]{Change the order of parameters for applyCast to take
  the value first and the cast second. -Jeremy}
\todo[inline]{The definition of values is too loose for a cast.
  You need to restrict the casts to just the inert ones. -Jeremy }
We conjecture that this CEK machines agree with the \lazyD{} cast calculus in 
\citet{siek2012interpretations}.
We will show in \hyperref{sec:hypercoercion} that this abstract machine is 
equivalent to an abstract machine that represents casts with hypercoercions.

@@

Transitive closure of reduction ($ \judgeDCreduceTrans{s}{s} $) and evaluation 
are 
standard.

\subsection{The \lazyUD{} cast calculus [Jeremy]}

The dynamic semantics of the \lazyUD{} cast calculus, which we define here as a
CEK machine, is similar to that of the \lazyD{} cast calculus
(Fig. \ref{machine-cekc}). The only difference is in the definition of
the $\mathit{applyCast}$ function, in which a cast whose source or
target is the unknown type $\star$ is always split into two casts that
go through an injectiable type, that is, a type in which all
sub-components are the unknown type, such as $\star \to \star$. The
definition of $\mathit{applyCast}$ for the \lazyUD{} cast calculus is given in
Fig.~\ref{fig:apply-Cast-UD}.

@@

\begin{lemma}[Weak Bisimulation between $ \mathcal{S}(\cdot) $ and $ 
	\mathcal{D} $]
	\label{thm:surely-monoidic-reduce}
	For all cast representations $ C $, 
	if $ C $ is surely \lazyD{} and monoidic,
	$ s_1 \in \mathcal{D} $,
	$ s_2 \in \mathcal{S}(C) $, and
	$ s_1 \approx s_2 $, 
	\begin{enumerate}
		\item If $ \judgeDCreduce{s_1}{s_3} $,
		then
		there exist $ s_5 \in \mathcal{D} $, $ s_4 \in \mathcal{S}(C) $,
		such that \begin{itemize}
			\item $ s_5 \approx s_4 $, and
			\item $ \judgeSreduce{C}{s_2}{s_4} $,
			\item $ \judgeDCreduceTrans{s_3}{s_5} $
		\end{itemize}
		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
		then
		there exist $ s_3 \in \mathcal{D} $,
		such that \begin{itemize}
			\item $ s_3 \approx s_4 $
			\item $ \judgeDCreduceTrans{s_1}{s_3} $
		\end{itemize}
	\end{enumerate}
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}

\begin{lemma}[Surely \lazyD\ and Monoidic Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-monoidic-eval}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{} and monoidic,
	\[
	\judgeDCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{lemma}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-reduce}.
\end{proof}

@@

\begin{theorem}[\lazyD{} hypercoercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeDCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval},
	Lemma~\ref{thm:hc-surely-lazyD}, and 
	Proposition~\ref{thm:hc-monoid}.
\end{proof}

@@

\begin{theorem}[Surely \lazyD\ Cast Representations Respect 
	$ \mathcal{D} $]
	\label{thm:surely-lazyD-correct}
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is surely 
	\lazyD{}
	\[
	\judgeDCeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{C}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:surely-monoidic-eval} and 
	Proposition~\ref{thm:surely-lazyD-eqv}.
\end{proof}
